Peer Assessments for Reproducible Research 
=============================================
The data was obtained from https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip
Time: May 17, 2015

First I loaded the csv file, and the dplyr package and ggplot2 package I will need later
```{r}
library(dplyr)
library(ggplot2)
activity<- read.csv("activity.csv")
summary(activity)
```
From the summary we can see that the date values are read in as factors. I would like to change them into dates. 
```{r}
activity$date<- as.Date(activity$date)
```

Then I would like to answer the first set of question: 
1, Total number of steps taken per day
2, Histogram of the total number of steps each day
3, Mean and Median of total number of steps taken perday. 

```{r}
totalstep<- activity %>% select(steps, date) %>% group_by(date) %>% summarize(total=sum(steps))
hist(totalstep$total,breaks=20, col="darkgrey", main="Total steps per day", xlab="Total steps")
Mean<- mean(totalstep$total, na.rm=T)
Median<- median(totalstep$total, na.rm=T)
```
Therefore the mean of the total number of daily steps is `r Mean`, and the median is `r Median`. 

Now is the second part of the assessment. The questions are:
1, Make a time series plot of 5-minute interval and the average steps
2, Identify the 5-min interval that has maximum number of steps. 
```{r}
intervalstep<- activity %>% group_by(interval) %>% summarize(average=mean(steps, na.rm=T)) 
plot(intervalstep$interval, intervalstep$average, type="l",col="red",lwd=2, main="Time series of average steps", xlab="5-min interval", ylab="Average steps")
Max<- max(intervalstep$average)
Maxinter<- intervalstep[intervalstep$average==Max,1]
```
Therefore, the interval that has maximum average step is `r Maxinter`. 

Now is the third part, imputing missing values. 
1, Report total number of missing values in the dataset
2, Fill in the missing value. I will use the mean of the corresponding interval. 
3, Make a new histogram and calculate mean and median. 
```{r}
Missingnum<- sum(is.na(activity))
Missing<- activity[is.na(activity$steps),]
Missing2<- merge(Missing, intervalstep)
Missing2<- Missing2[,-2]
names(Missing2)[3]='steps'
activitynew<- rbind(na.omit(activity), Missing2)
total2<- activitynew %>% select(steps, date) %>% group_by(date) %>% summarize(total=sum(steps))
hist(total2$total, breaks=20, col="blue", main="Histogram of total daily steps (with imputed data)", xlab="Total daily steps")
Mean2<- mean(total2$total)
Median2<- median(total2$total)
```
Before imputing missing data, the total number of missing value is `r Missingnum`. 
After imputation, we can see that the new mean is `r Mean2`, and the new median is `r Median2`. The mean stayed the same as before, but the median changed slightly. The histogram, however, looks quite different from the previous one. The frequency of the bar for 10000~11000 steps significantly increased. This is because that I used the average step number for a certain interval to impute the missing data, so the average value is 'enhanced' and the frequency increased 

Now come the last part of the assessment. The challenge is to make a time series plot for average steps on weekdays and weekend. 
```{r}
weekday<- weekdays(activitynew$date)
weekend<- weekday %in% c("Saturday", "Sunday")
Cat<- rep("Weekday", length(weekday))
Cat[weekend]="Weekend"
activitynew2<- cbind(activitynew, Cat)
interval2<- activitynew2 %>% group_by(Cat, interval) %>% summarize(average=mean(steps))
g<- qplot(interval, average, data=interval2, facets=Cat~., geom="line")
g+labs(title="Average steps across interval", x="Interval", y="Number of steps")
```





